---
title: "STAT 542 Project 4"
author: "Mehrdad Mohammadi - David Kim"
date: "12/12/2021"
output:
  html_document: default
  fontsize: 10pt
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{amsthm}
- \usepackage{amsmath}
- \usepackage{amsfonts}
- \usepackage{amscd}
- \usepackage{amssymb}
- \usepackage{xcolor}
- \usepackage{bbm}
---

---
\allowdisplaybreaks

\newcommand{\R}{\mathbb{R}}
\newcommand{\p}{\mathbb{P}}
\newcommand{\Proj}{\textbf{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Hcal}{\mathcal{H}}
\newcommand{\rootn}{\sqrt{n}}
\newcommand{\pibf}{\bm{\pi}}
\newcommand{\1}{\mathbb{1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



# Table of Contents
1.  [Part I:  Pre-Analysis ](#Part I:  Pre-Analysis)\
2.  [Part II: Recommender System  by Genre ](#Part II: Recommender System  by Genre )\
2.1 [Recommendation Schemes : top  most popular movies](##Recommendation Schemes : top  most popular movies in the genre)\
2.2 [Recommendation Schemes : top  highly-rated ](##Recommendation Schemes : top  highly-rated )\
3.  [Part III: Collaborative Recommender System](#Part III: Collaborative Recommender System)\
3.1 [Recommendation Schemes: UBCF](##Recommendation Schemes: USER BASED COLLABORATIVE RECOMMENDER)\
3.2 [Recommendation Schemes: IBCF](##Recommendation Schemes: IBCF)\
3.3 [Evaluation](##Evaluation)

```{r}
# Pakcages
library(devtools)
library(recommenderlab)
library(ggplot2)
library(data.table)
library(Matrix)
library(dplyr)
library(DT)
library(reshape2)
```

# Part I:  Pre-Analysis 
## Data
```{r}
#read data
myurl  = "https://liangfgithub.github.io/MovieData/"
# Ratings
ratings = read.csv(paste0(myurl, 'ratings.dat?raw=true'), 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
ratings$Timestamp = NULL
# Movie
movies = readLines(paste0(myurl, 'movies.dat?raw=true'))
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)
movies$Title = iconv(movies$Title, "latin1", "UTF-8")

small_image_url = "https://liangfgithub.github.io/MovieImages/"
movies$image_url = sapply(movies$MovieID, 
                          function(x) paste0(small_image_url, x, '.jpg?raw=true'))
# User
users = read.csv(paste0(myurl, 'users.dat?raw=true'),
                 sep = ':', header = FALSE)
users = users[, -c(2,4,6,8)] # skip columns
colnames(users) = c('UserID', 'Gender', 'Age', 'Occupation', 'Zip-code')

```

# Part II: Recommender System  by Genre 

## Genres in data
```{r}
# Genres in data
#Splitting genres from movie table
genre     <- as.data.frame(movies$Genres, stringsAsFactors=FALSE)
#Classifies each movie into different categories
genre_new <- as.data.frame(tstrsplit(genre[,1], '[|]', type.convert=TRUE), stringsAsFactors=FALSE)

#getting unique list of genres
genre_list = c()
for (j in 1:6) {
genre_list = c(genre_list,unique(genre_new[,j]))
}
genre_list <- c(unique(na.omit(genre_list)))
genre_matrix = matrix(0, nrow(movies), length(genre_list))
for(i in 1:nrow(genre_new)){
  genre_matrix[i,genre_list %in% genre_new[i,]]=1
}
colnames(genre_matrix) = genre_list
remove("genre_new", "genre")
genre_list
```

## Recommendation Schemes : top  most popular movies in the genre
The movies that are highly rated i.e. at least 1000 users have rated the movie with ratings of 4 stars or 5 stars in the selected genre are recommended.
```{r}
selected_genre = "Action"

 selectedMoviesByGenre = subset(movies, grepl(selected_genre, movies$Genres, fixed = TRUE), select = c(MovieID, Title))
      
      selectedMoviesByGenre$MovieID = as.integer(selectedMoviesByGenre$MovieID)
      # Select movies with ratings >= 4
      selectedMoviesByRating = subset(ratings, Rating>=4, select = c(MovieID, Rating))
      system1Result = selectedMoviesByRating %>% inner_join(selectedMoviesByGenre, by= "MovieID")
      rating_summary = system1Result %>% group_by(Rating) %>% summarise(rating_count=n())
      
      final_selected_movies = c()
      if (rating_summary$rating_count[1] > 1000 || rating_summary$rating_count[2] > 1000) {
        final_selected_movies = head(subset(system1Result[sample(nrow(system1Result), 50),], !duplicated(MovieID)), 10)
      }
      
      final_selected_movies
```

## Recommendation Schemes : top  highly-rated 

The movies that are highly rated i.e. at least 1000 users have rated the movie with ratings of 5 stars in the selected genre are recommended.

```{r}
selected_genre = "Action"

 selectedMoviesByGenre = subset(movies, grepl(selected_genre, movies$Genres, fixed = TRUE), select = c(MovieID, Title))
      
      selectedMoviesByGenre$MovieID = as.integer(selectedMoviesByGenre$MovieID)
      # Select movies with ratings == 5
      selectedMoviesByRating = subset(ratings, Rating==5, select = c(MovieID, Rating))
      system1Result = selectedMoviesByRating %>% inner_join(selectedMoviesByGenre, by= "MovieID")
      rating_summary = system1Result %>% group_by(Rating) %>% summarise(rating_count=n())
      
      final_selected_movies = c()
      if (rating_summary$rating_count[1] > 1000 || rating_summary$rating_count[2] > 1000) {
        final_selected_movies = head(subset(system1Result[sample(nrow(system1Result), 50),], !duplicated(MovieID)), 10)
      }
      
      final_selected_movies
```

# Part III: Collaborative Recommender System 


```{r}
# 2.1.Training Test Split

set.seed(8439)
# create a utility matrix stored as a sparse matrix.
i = paste0('u', ratings$UserID)
j = paste0('m', ratings$MovieID)
x = ratings$Rating
tmp = data.frame(i, j, x, stringsAsFactors = T)
Rmat = sparseMatrix(as.integer(tmp$i), as.integer(tmp$j), x = tmp$x)
rownames(Rmat) = levels(tmp$i)
colnames(Rmat) = levels(tmp$j)
Rmat = new('realRatingMatrix', data = Rmat)
e <- evaluationScheme(Rmat, method="split", train=0.8, given=15, goodRating=5)
```

## Recommendation Schemes: USER BASED COLLABORATIVE RECOMMENDER
The assumption is that users with similar preferences will rate items similarly. Thus missing ratings for a user can be predicted by first finding a neighborhood of similar users and then aggregate the ratings of these users to form a prediction.

The neighborhood is defined in terms of similarity between users, either by taking a given number of most similar users (k nearest neighbors)or all users within a given similarity threshold.  Our similarity measures is $\textbf{Cosine similarity}$ and $\textbf{we used default values on the similarity measure}$. As default values shown in the codes below for these parameter.
```{r}
##### USER BASED COLLABORATIVE RECOMMENDER ########
recommenderRegistry$get_entries(dataType="realRatingMatrix")$UBCF_realRatingMatrix

# Train a recommender system using R package
rec_UBCF = Recommender(getData(e, "train"), "UBCF",
                        parameter = list(normalize = 'Z-score', method = 'Cosine', nn = 25))
# predict ratings for the users
recom_UBCF = predict(rec_UBCF,getData(e, "known"), type="ratings")

## save and read a recommender model
saveRDS(rec_UBCF, file = "rec_UBCF.rds")
rec_UBCF2 <- readRDS("rec_UBCF.rds")

# A sample of predictions ratings for the users
as(recom_UBCF, 'matrix')[1:3, 1:10]
```


## Recommendation Schemes: ITEM BASED COLLABORATIVE RECOMMENDER
Item-based CF is a model-based approach which produces recommenda-
tions based on the relationship between items inferred from the rating matrix. The assumption behind this approach is that users will prefer items that are similar to other items they like. The model-building step consists of calculating a similarity matrix containing all item-to-
item similarities using a given similarity measure. Popular are again Pearson correlation and Cosine similarity. To make a recommendation based on the model we use the similarities to calculate a weighted
sum of the user's ratings for related items. 
Similar to user-based recommender algorithms, user-bias can be reduced by first normalizing
the user-item rating matrix before computing the item-to-item similarity matrix.

```{r}
##### IBCF RECOMMENDER ########
recommenderRegistry$get_entries(dataType="realRatingMatrix")$IBCF_realRatingMatrix
# Train a recommender system using R package
rec_IBCF = Recommender(getData(e, "train"), "IBCF",
                       parameter = list(normalize = 'Z-score', method = 'Cosine', nn = 25))
# predict ratings for the users
recom_IBCF = predict(rec_IBCF,getData(e, "known"), type="ratings")

## save and read a recommender model
saveRDS(rec_IBCF, file = "rec_IBCF.rds")
rec_IBCF2 <- readRDS("rec_IBCF.rds")
# A sample of predictions ratings for the users
as(recom_IBCF, 'matrix')[1:3, 1:10]
```


## Evaluation
We can calculate the error between the prediction and the unknown part of the test data.
```{r}
error <- rbind(UBCF = calcPredictionAccuracy(recom_UBCF, getData(e, "unknown")),
               IBCF = calcPredictionAccuracy(recom_IBCF, getData(e, "unknown")))
error
```
